Assessment of the KLGO Coastal Waterbird Survey Monitoring Program

### 2019-05-08 analysis/paper/paper.Rmd
site accessibility attempt
Lots of changes:  clarity within scripts, new scripts in paper (survey access, species occurrence)

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = TRUE, # set to FALSE when done
  message = TRUE, # set to FALSE when doen
  echo = FALSE,
  comment = "#>",
  dev = c('png','pdf'),
  pdf.options(encoding = "ISOLatin9.enc"),
  fig.path = "../figures/",
  error = FALSE
)

@@

```
# NOTES To Madeleine
- see edits in survey_duration.R to aid interpretation by the viewer (e.g., changes in x axis tick locations, etc.)
- documentation in scripts
- **NEXT Priority Milestones**: note - all references to line numbers are prior to any other edits!
- finish graphs of first & last observation per year for each target species (around line 386);**ultimately** add a fitted linear trend line to each of those (we need to account for frequency of surveying, so some subtle effort things here, but can resolve that later);
- graphs of species occurrence (around line 407) for each of target species; 
- covariate distribution graphs (lots of univariate & bivariate things; around line 534);
- plot of total # spp sighted each season across years (around line 600)

@@

#NOTE: working directory when knitting markup is /analysis/paper. So relative pathway needs to be ../scripts etc.
source("../scripts/data_retrievalCSV.R")

@@

> SECOND graph: [THIS IS SECONDARY PRIORITY] x axis is Time of Day, perhaps broken in to 1 or 2 hour blocks , y axis is ....this is a tough one so totally open to your ideas. How about y axis is % [0, 1.0] and for each year there is a different curve showing the % of planned surveys (as defined above) that were successfully completed that season within that time block; facet by site. So we are seeing whethere there are consistently problematic times of day at each site. 
```{r SiteAccess, eval=TRUE, cache = FALSE, include = FALSE}
# STILL NEED TO CREATE
# source("../scripts/survey_access.R", echo=FALSE)
```

```{r Fig-SiteAccess, echo=FALSE, eval=TRUE, cache=TRUE, fig.show= 'asis',fig.height=20, fig.width=12, fig.cap="",warning= FALSE}
# print(plot_access)
```

@@

```{r Fig-SurveyEffort, echo=FALSE, eval=TRUE, cache=TRUE, fig.show= 'asis', fig.height=15, fig.width=12,fig.cap="",warning= FALSE}
print(plot_duration)
```

@@

> MADELEINE: add a graphic of spp frequency of occurrence (maybe some panel-rich figure, one panel per spp, showing annual average percentage of daily observations in which spp was sighted - as line-connected points?- ordered from most frequently observed spp to least frequently. (e.g., use facets). x axis is Day of Season, y axis is % as decimal [0, 1.0], curves for each year; facet by species. LET"S DISCUSS - Think about appropriate estimator so annual value accounts for variation in survey effort across years.) likely needs to be a full page figure (e.g., we'll need to set  fig.width and fig.height). TAKE A STAB AT THIS BUT DON'T get lost trying to complete all the refinements. Really we should build up to this.

```{r EDA_AllSpp_Frequency, eval=TRUE, cache = FALSE, include = FALSE, fig.height=8, fig.width=6}
# lets discuss what I'm thinking of. It is of secondary importance, so don't want this 
# to consume tons of your time.

#source("../scripts/survey_timing_EDA.R", echo=FALSE)

@@

```{r Fig-AllSppFreq, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis', fig.cap="blah blah about consistency in target species occurrence. Maybe something about new spp showing up?"}

#print(Plot1)
```

@@

```{r SiteSurveyCompletion, eval=TRUE, cache = FALSE, include = FALSE}
# dotplot of survey occurrences by unit and whether there were observations
source("../scripts/survey_sites.R", echo=FALSE)
```

@@

```{r Fig-SiteSurveyCompletion, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis',fig.height=10, fig.width=18, fig.cap="Need to add caption.",warning=FALSE}

@@

```{r EDA_SurveyTiming, eval=TRUE, cache = FALSE, include = FALSE}
# dotplot of survey times by year
source("../scripts/survey_timing_EDA.R", echo=FALSE)

```{r Fig-SurveyTiming1, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis',fig.height=10, fig.width=12, fig.cap="The annual dates of the first and last survey have both varied by over a month across the years."}

@@

```{r SurveyDates, eval=TRUE, cache = FALSE, include = FALSE}
# scatterplot of species' first and sightings compared to survey season start
source("../scripts/survey_dates.R", echo=FALSE)
```

```{r Fig-SurveyDates1, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis', fig.cap="The first sighting of each species each year.",fig.height=20, fig.width=12, warning=FALSE}
print(plot_timing_1)
```

@@

```{r Fig-SurveyDates2, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis', fig.height=20, fig.width=12, fig.cap="The last sighting of each species each year.", warning=FALSE}
print(plot_timing_2)
```

@@

>>>>

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = TRUE, # set to FALSE when done
  message = TRUE, # set to FALSE when done
  echo = FALSE,
  comment = "#>",
  dev = c('png','pdf'),
  pdf.options(encoding = "ISOLatin9.enc"),
  fig.path = "../figures/",
  error = FALSE
)

@@

```
# NOTES To Madeleine
- see edits in survey_duration.R to aid interpretation by the viewer (e.g., changes in x axis tick locations, etc.) $\checkmark$
- documentation in scripts $\checkmark$
- **NEXT Priority Milestones**: note - all references to line numbers are prior to any other edits!
- finish graphs of first & last observation per year for each target species (around line 386);**ultimately** add a fitted linear trend line to each of those (we need to account for frequency of surveying, so some subtle effort things here, but can resolve that later); $\checkmark$
- graphs of species occurrence (around line 407) for each of target species; $\checkmark$
- covariate distribution graphs (lots of univariate & bivariate things; around line 534);
- plot of total # spp sighted each season across years (around line 600)

@@

#NOTE: working directory when knitting markup is /analysis/paper. So relative pathway needs to be ../scripts etc.

source("../scripts/data_retrievalCSV.R")

@@

> SECOND graph: [THIS IS SECONDARY PRIORITY] x axis is Time of Day, perhaps broken in to 1 or 2 hour blocks , y axis is ....this is a tough one so totally open to your ideas. How about y axis is % [0, 1.0] and for each year there is a different curve showing the % of planned surveys (as defined above) that were successfully completed that season within that time block; facet by site. So we are seeing whether there are consistently problematic times of day at each site. 
> JOEL: some updates on this! for now, times are in hour blocks. There can't be lines/curves for year since they are unreadable (up/down). I also tried to use color as a factor but then the graphs are crowded (so many years). So I have 2 alternatives, I don't really like either but some are more readable than others. See below.

```{r SiteAccess, eval=TRUE, cache = FALSE}

source("../scripts/survey_access.R", echo=FALSE)
```

```{r Fig-SiteAccess1, echo=FALSE, eval=TRUE, cache=TRUE, fig.show= 'asis',fig.height=20, fig.width=12, fig.cap="",warning= FALSE}

print(plot_success)
```

```{r Fig-SiteAccess2, echo=FALSE, eval=TRUE, cache=TRUE, fig.show= 'asis',fig.height=15, fig.width=12, fig.cap="",warning= FALSE}

print(plot_success3)
```

@@

```{r Fig-SurveyEffort, echo=FALSE, eval=TRUE, cache=TRUE, fig.show= 'asis', fig.height=12, fig.width=12, fig.cap="",warning= FALSE}

print(plot_duration)
```

@@

> MADELEINE: add a graphic of spp frequency of occurrence (maybe some panel-rich figure, one panel per spp, showing annual average percentage of daily observations in which spp was sighted - as line-connected points?- ordered from most frequently observed spp to least frequently. (e.g., use facets). x axis is Day of Season, y axis is % as decimal [0, 1.0], curves for each year; facet by species.
> JOEL: I did some graphs for species occurrence that fit the description but not sure if they were what you were looking for? They do show species occurrence relative to others; as annual/daily percentage of daily observations in which spp was sighted (percentage calculated by annual/daily obs per species over all total obs - is this what you meant? let me know).
Lines didn't work though (lots of ups and downs)

```{r EDA_AllSpp_Frequency, eval=TRUE, cache = FALSE, include = FALSE, fig.height=8, fig.width=6}

source("../scripts/occurrence_freq.R", echo=FALSE)

@@

```{r Fig-AllSppFreq1, echo=FALSE, eval=TRUE, warning=FALSE, cache=FALSE, fig.show= 'asis', fig.height=20, fig.width=12, fig.cap="blah blah about consistency in target species occurrence. Maybe something about new spp showing up?"}

print(plot_occurrenceDay)
```

```{r Fig-AllSppFreq2, echo=FALSE, eval=TRUE, warning=FALSE, cache=FALSE, fig.show= 'asis', fig.height=25, fig.width=10, fig.cap="blah blah about consistency in target species occurrence. Maybe something about new spp showing up?"}

print(plot_occurrenceYear)
```

@@

```{r SiteSurveyCompletion, eval=TRUE, cache = FALSE, include = FALSE}
# dotplot of survey occurrences by unit and whether there were observations

source("../scripts/survey_sites.R", echo=FALSE)
```

@@

```{r Fig-SiteSurveyCompletion, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis',fig.height=15, fig.width=18, fig.cap="Need to add caption.",warning=FALSE}

@@

```{r EDA_SurveyTiming, eval=TRUE, cache = FALSE, include = FALSE}
# dotplot of survey times by year

source("../scripts/survey_timing_EDA.R", echo=FALSE)
```

```{r Fig-SurveyTiming1, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis',fig.height=8, fig.width=12, fig.cap="The annual dates of the first and last survey have both varied by over a month across the years."}

@@

```{r SurveyDates, eval=TRUE, cache = FALSE, include = FALSE}

# scatterplot of species' first and last sightings compared to survey season start
source("../scripts/survey_dates.R", echo=FALSE)
```

```{r Fig-SurveyDates1, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis', fig.cap="The first sighting of each species each year.",fig.height=20, fig.width=12, warning=FALSE}

print(plot_timing_1)
```

@@

```{r Fig-SurveyDates2, echo=FALSE, eval=TRUE, cache=FALSE, fig.show= 'asis', fig.height=20, fig.width=12, fig.cap="The last sighting of each species each year.", warning=FALSE}

print(plot_timing_2)
```

@@

MADELEINE
